Describing the C-Space of Asynchronous Periodic Task Systems using Definitive Idle Times

### 2013-07-30 dit-paper_ieee.tex
Changes after discussion w/ LG

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

@@

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Describing the C-Space of Asynchronous Periodic Task Systems using
Definitive Idle Times}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Thomas Chapeaux\\ and Paul Rodriguez}
\IEEEauthorblockA{Universit\'{e} Libre de Bruxelles\\
Email: tchapeau@ulb.ac.be, paurodri@ulb.ac.be}
\and
\IEEEauthorblockN{Laurent George\\ and Jean-Fran\c{c}ois Hermant}
\IEEEauthorblockA{ECE Paris\\
Email}
}
% make the title area
\maketitle

\begin{abstract}
%\boldmath
Sensitivity analysis for real-time systems asks for efficient methods to determine feasibility
after small changes of execution times, such as when porting a system to a different platform.
The C-space, presented in \cite{bini2004schedulability,george2009characterization}, describes the
set of execution times for which a system is feasible. This paper extends previous works to
generalize the description of the C-space for Earliest Deadline First scheduling to asynchronous
systems and non-pairwise coprime periods. The feasibility gain offered by offsets is then
expressed as the volume ratio between the C-space of an asynchronous system and its corresponding
synchronous system.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

@@

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, the smallest description is preferred as it can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

  \subsection{Removing redundancy}
    \label{sct:removeRedundancy}

@@

  For synchronous systems in which the execution times are not known, the first DIT $t_d$
  is our ``best guess'' at an upper limit for the end of the busy period. As it was sufficient
  to check the DBF test in intervals $[0, t]$ before the end of the busy period, it is
  slightly more pessimistic but still correct to use the following condition:
  \[
  	\begin{array}{c}
  		\dbf{0,t} \leqslant t \; \forall t \leqslant t_d \\
  		\iff \text{The system is feasible}
  	\end{array}
  \]

@@

  \subsection{Our work}
    In this paper, we extend the work of \cite{george2009characterization} to the
    asynchronous case. In section \ref{sct:asyncDIT}, we present an extension of
    the concept of DIT adapted to the asynchronous case called first periodic DIT
    ($t_d$) along with an algorithm to determine its existence and value for any system,
    then we show the influence of some factors on its existence. In Section
    \ref{sct:asyncCspace}, we show how to describe the C-space of an asynchronous
    system based on the DBF test, we show why the interval $[t_d, t_d + H]$ is
    sometimes sufficient to consider and we give a numerical example.
    Finally, in section ???, we present a simulation where we find that........

@@

	\subsection{Removing redundancy}

@@

Fig. \ref{fig:cspaceComp} shows the two C-spaces in the plane.

@@

\section{Experimental C-space size gain of asynchronous systems}
	One of the advantages of asynchronous systems over synchronous systems in the
	context of periodic tasks is that the critical instants of asynchronous systems
	cannot be more constraining for the C-space than synchronous instants. For this
	reason, it has long been known that modifying task offsets when possible during
	the design of a real-time system is a useful trick to fit a heavy but periodic task set
	into a given platform.

	In Fig.~\ref{fig:sizeRatio} a number of feasible asynchronous task sets of
	three tasks without synchronous instants were generated. The size of their
	C-space was calculated and compared with the size of the C-space of the same
	task sets with all task offsets set to zero. In other words, this figure
	evaluates the feasibility gain of adding random offsets to a synchronous task
	set. It is assumed that a more advanced strategy to assign these offsets
	would yield a greater difference.

@@

	The ratio between the size of the synchronous and asynchronous C-spaces of
	these systems roughly range from 0.95 in the worst case (high utilization, high
	CDF) to 0.85 in the best case (low utilization, low CDF). Note that the value
	of the CDF has a greater influence on this ratio than the system's total
	utilization, but less so when the CDF is lower. It is also important to note
	that even though a 0.95 ratio may not seem much, the synchronous C-spaces of
	the systems contain many dominated WCET vectors, and only those that are
	pareto-optimal (i.e. non-dominated) actually matter to the system's engineer.
	Using the asynchronous C-space instead will relax some constraints and therefore add new acceptable
	solutions that will dominate solutions that were previous non-dominated,
	increasing the quality of the WCET pareto-front.

@@

In this paper, we presented a general approach to define the C-space of a constrained periodic task set for EDF on uniform platform. Possible future works includes an extension to multiprocessor platforms or including preemption costs, in which case EDF is no longer an optimal algorithm, meaning that another algorithm could be studied.

(...)

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \nocite{*}
\bibliographystyle{IEEEtran}
\bibliography{dit-paper}

>>>>

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

@@

@@

\begin{document}
\title{Describing the C-Space of Asynchronous Periodic Task Systems using
Definitive Idle Times}

\author{
\IEEEauthorblockN{Thomas Chapeaux\\ and Paul Rodriguez}
\IEEEauthorblockA{Universit\'{e} Libre de Bruxelles\\
Email: tchapeau@ulb.ac.be, paurodri@ulb.ac.be}
\and
\IEEEauthorblockN{Laurent George\\ and ???}
\IEEEauthorblockA{ECE Paris\\
Email}
}
\maketitle

\begin{abstract}
Sensitivity analysis for real-time systems asks for efficient methods to determine feasibility
after small changes of execution times, such as when porting a system to a different platform.
The C-space, presented in \cite{bini2004schedulability,george2009characterization}, describes the
set of execution times for which a system is feasible. This paper extends previous works to
generalize the description of the C-space for Earliest Deadline First scheduling to asynchronous
systems and non-pairwise coprime periods. The feasibility gain offered by offsets is then
expressed as the volume ratio between the C-space of an asynchronous system and its corresponding
synchronous system.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

\section*{TODO}

\begin{itemize}
	\item changer CDF
	\item Expliquer simulation plus en d\'etail (GLPK)
	\item Donner un algo pour CRP
	\item fix bibliography
\end{itemize}

@@

      However, this description contains an infinite number of constraints, most of
      which are redundant. While this does not harm the precision of the
      test, a smaller description is preferred as it can greatly
      reduce the computation time of practical feasibility tests. In Section~\ref{sct:removeRedundancy} a method to reduce the size of this set of constraints is given.

  \subsection{Removing redundant constraints}
    \label{sct:removeRedundancy}

@@

  For synchronous systems in which the execution times are not known, the first DIT $t_d$
  is our ``best guess'' at an upper limit on the end of the busy period. As it was sufficient
  to check the DBF test in intervals $[0, t]$ before the end of the busy period, it is
  slightly more pessimistic but still correct to use the following condition:
  \[
  	\begin{array}{c}
  		\dbf{0,t} \leqslant t \; \forall t \leqslant t_d \\
  		\iff \text{The system is feasible}
  	\end{array}
  \]

@@

  \subsection{Our work}
    In this paper, we extend the work of \cite{george2009characterization} to the
    asynchronous case. In section \ref{sct:asyncDIT}, we present an extension of
    the concept of DIT adapted to the asynchronous case called first periodic DIT
    ($t_d$) along with an algorithm to determine its existence and value for any system,
    then we show the influence of some factors on its existence. In Section
    \ref{sct:asyncCspace}, we show how to describe the C-space of an asynchronous
    system based on the DBF test, we show why the interval $[t_d, t_d + H]$ is
    sometimes sufficient to consider and we give a numerical example.
    Finally, in section \ref{sct:expCspaceGain}, we use a simulation to quantify the feasibility
    gain of asynchronous system in terms of volume of the C-space.

@@

	\subsection{Removing redundant constraints}

		The first two following sections contain well-known results of feasibility analysis.
		The third one is a new result based on the FPDIT and the last section contains a general
		but heavy method for removing constraints redundancies.

@@

Fig. \ref{fig:cspaceComp} shows the two C-spaces in the plane. The dark grey region is the C-space of the synchronous system and the light grey region is the surface gain given by the asynchronous system.

@@

\section{Experimental C-space size gain of asynchronous systems}
\label{sct:expCspaceGain}
	One of the advantages of asynchronous systems over synchronous systems in the
	context of periodic tasks is that the critical instants of asynchronous systems
	are at most as constraining for the C-space as a synchronous instant. For this
	reason, modifying task offsets when possible during
	the design of a real-time system is a useful trick to fit a heavy periodic task set
	into a given platform.

	In Fig.~\ref{fig:sizeRatio} a number of feasible asynchronous task sets of
	three tasks without synchronous instants were generated. The size of their
	C-space (number of valid WCET vectors) was calculated and compared with that of the same
	task sets with all offsets set to zero. In other words, this figure
	evaluates the feasibility gain of adding random offsets to a synchronous task
	set. It is assumed that a more advanced strategy to assign these offsets
	would yield a greater difference.

@@

	The ratio between the size of the synchronous and asynchronous C-spaces of
	these systems roughly range from 0.95 in the worst case (high utilization, high
	CDF) to 0.85 in the best case (low utilization, low CDF). Note that the value
	of the CDF has a greater influence on this ratio than the system's total
	utilization, but less so when the CDF is lower. It is also important to note
	that even though a 0.95 ratio may not seem much, the synchronous C-spaces of
	the systems contain many dominated WCET vectors, and only those that are
	pareto-optimal (i.e. non-dominated) actually matter to the designer of the system.
	Using the asynchronous C-space instead will relax some constraints and therefore add new acceptable
	solutions that will dominate solutions that were previously non-dominated,
	increasing the quality of the WCET pareto-front.

@@

In this paper, we presented a general approach to define the C-space of a constrained periodic task set for EDF on uniform platform. Possible future works includes an extension to the study of systems with preemption costs or running on multiprocessor platforms, as the value of the DIT does not depend on preemptions or the number of processors.

Future works could also study the shape of the C-space and how it is affected by some factors. An interesting path would be the effect of the CDF on the number of constraints describing the C-space, as we know that in the implicit case this description is simple ($U_{tot} \leqslant 1$). While we quantified the feasibility gain by looking at the ratio between sizes of C-space, another approach would be to look at the ratio between the minimal distance of the pareto front of the C-space, which would give a lower bound of the gain given by the offsets.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \nocite{*}
\bibliographystyle{plain}
\bibliography{dit-paper}
